---
title: "Survived Titanic Passenger Prediction with Logistic Model and KNN"
author: "Rahma Fairuz Rania"
date: '2022-07-09'
output: 
 html_document:
   toc: true
   toc_float: true
   highlight: zenburn
   df_print: paged
   theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Case : From data titanic passenger, we want to know does passenger still alive or not. Data can be downloaded here https://www.kaggle.com/c/titanic

# Import Data

```{r}
titanic_train <- read.csv('train.csv')
titanic_test <- read.csv('test.csv')
head(titanic_train)
#head(titanic_test)
```

# Data Wrangling

To make easier, combine data train and test into 1 dataframe

```{r, warning=FALSE, message=FALSE}
library(dplyr)
titanic_test$Survived <- NA

titanic <- rbind(titanic_train, titanic_test) %>% 
    mutate(Survived = as.factor(Survived))
```

- Check structure of data

```{r}
str(titanic)
```

Here are some information about columns in wholesale data
* `PassengerId`: Id number of passenger
* `Survived` : Passenger Survival or Not
* `Pclass` :  A proxy for socio-economic status (SES). 1st = Upper, 2nd = Middle, 3rd = Lower
* `Name` : Name of Passenger
* `Sex` : Sex of Passenger 
* `Age` : Age of Passenger
* `SibSp` : Sibling = brother, sister, stepbrother, stepsister; Spouse = husband, wife (mistresses and fiancés were ignored)
* `Parch` :  Parent = mother, father; Child = daughter, son, stepdaughter, stepson; Some children travelled only with a nanny, therefore parch=0 for them.
* `Ticket` : Ticket number
* `Fare` : Passenger fare
* `Cabin` : Cabin number
* `Embarked` : Port of Embarkation 

- Check proportion 

```{r}
prop.table(table(titanic$Survived))
```
Our data has balance enough.

- Balancing class proportion with downsampling

```{r}
# downsampling
RNGkind(sample.kind = "Rounding")
set.seed(100)
library(caret)
titanic <- downSample(x = titanic %>% 
                           select(-Survived),
                          y = titanic$Survived,
                         yname = "Survived") #nama kolom target

head(titanic)
```

```{r}
#cek proporsi diab_train
prop.table(table(titanic$Survived))
```
Our data has balance proportion

- Check missing value

```{r}
anyNA(titanic)
```

```{r}
# remove missing value
titanic <- titanic %>% na.omit()
anyNA(titanic)
```

# Cross Validation

Give proportion to data train 80% and 20% to data test. Proportion data train bigger than data test because we want our model learn more.

```{r, warning=FALSE, message=FALSE}
RNGkind(sample.kind = "Rounding")
set.seed(100)

# index sampling
index <- sample(x = nrow(titanic), 
                size = nrow(titanic)*0.8)

# splitting
titanic_train <- titanic[index,] 
titanic_test <- titanic[-index,]
```

- Check data train proportion

```{r}
prop.table(table(titanic_train$Survived))
```
Our data class proportion is balance enough.

# Data Pre-processing

### Logistic Regression

- Build model with predictors Pclass, Sex, Age, SibSp, Parch, and Embarked

```{r, warning=FALSE, message=FALSE}
model_logistic <- glm(formula = Survived ~ Pclass + Sex + Age + SibSp + Parch + Embarked,
                   data = titanic,
                   family = "binomial")
```


- Summary model

```{r}
summary(model_logistic)
```

Log of odds value from logistic model can't be interpreted, we should convert it into Odds.

```{r}
exp(model_logistic$coefficients)
```

We can explain how each variable significant to prediction. According to model summary, passenger in Pclass 2 2.81 times more likely survived. Passenger with parent or children 9.56 times more likely survived.

- Feature Selection

Choose significant variable to our model

```{r}
model_step <- step(model_logistic, direction = 'backward', trace = F)
summary(model_step)
```

- Predict data test 

```{r}
titanic_test$predict_value <- predict(object = model_step,
                      newdata = titanic_test,
                      type = "response")
```

Predict result type probability, transform it into label according to target variable(Survived) 

```{r}
unique(titanic_test$Survived)
```

```{r}
titanic_test$pred_label <- ifelse(test = titanic_test$predict_value > 0.5,
                                yes = "1",
                                no = "0")

titanic_test$pred_label <- as.factor(titanic_test$pred_label)
head(titanic_test)
```

- Model evaluation with confusionMatrix from library caret

```{r}
library(caret)
confusionMatrix(data = titanic_test$pred_label,
                reference = titanic_test$Survived,
                positive = "1")
```
Logistic model has 78% accuracy, 78% sensitivity, 77% specificity, and 83% precision. Next, we want to compare this model with another model (K-Nearest Neighbor) to see which model better in predict.

### K-Nearest Neighbor

- Data cleaning, we just use numerical column

```{r}
titanic_clean <- titanic %>% select(-c(Name, Sex, Ticket, Fare, Cabin, Embarked))
```

- Check class proportion

```{r}
prop.table(table(titanic_clean$Survived))
```
- Cross validation

```{r, warning=FALSE, message=FALSE}
RNGkind(sample.kind = "Rounding")
set.seed(257)

# data train 80%
index <- sample(x = nrow(titanic_clean), 
                size = nrow(titanic_clean)*0.8)

# splitting
titanic_train_for_knn <- titanic_clean[index, ]
titanic_test_for_knn <-  titanic_clean[-index, ]

prop.table(table(titanic_train_for_knn$Survived)) # recheck class proportion
```
Our data is balance enough

- Before we build KNN model, we should split target and predictor in both data train and data test.

```{r}
# predictor train
titanic_train_predictor <- titanic_train_for_knn %>% 
  select(-Survived)

# target train
titanic_train_target <- titanic_train_for_knn %>% 
  pull(Survived) 


# predictor test
titanic_test_predictor <- titanic_test_for_knn %>% 
  select(-Survived)

# taget test
titanic_test_target <- titanic_test_for_knn %>% 
  pull(Survived)
```

- Scale data train predictor for range standarization of variable predictor

```{r}
titanic_train_predictor_scale <- titanic_train_predictor %>% 
  scale()
```

```{r}
titanic_test_predictor_scale <- titanic_test_predictor %>% 
  scale(center = attr(titanic_train_predictor_scale,"scaled:center"),
        scale = attr(titanic_train_predictor_scale,"scaled:scale"))
```

- Predict with knn() from library class

Finding k optimum

```{r}
sqrt(nrow(titanic_train_predictor))
```
K optimum is 22

```{r, warning=FALSE, message=FALSE}
library(class)
titanic_pred <- knn(train = titanic_train_predictor_scale,
                   test = titanic_test_predictor_scale,
                   cl = titanic_train_target,
                   k = 20)
titanic_pred
```

- Model evaluation with confusionMatrix from library caret

```{r}
library(caret)
confusionMatrix(data = titanic_pred, reference = titanic_test_target, positive = '1')
```
KNN model has 72% accuracy, 72% sensitivity, 73% specificity, and 77% precision.

# Conclusion

Logistic model has 
```{r}

```

- 78% accuracy
- 78% sensitivity
- 77% specificity
- 83% precision


KNN model has
```{r}

```

- 72% accuracy
- 72% sensitivity
- 73% specificity
- 77% precision

> In this case, we focused on recall or sensitivity. So, Logistic Model better than KNN Model to predict passenger who survived from titanic sink disaster.